\bf Creating optimal conditions for reproducible data analysis in R with \texttt{fertile}

### 2019-11-24 fertile.Rmd
start of extended abstract draft

@@

Practicing reproducible research is key for advancing our scientific knowledge. One reason for this is that reproducible research encourages transparency. Providing one's audience with well organized, reproducible code and results allows readers to understand the steps taken to generate findings from raw data and determine for themselves whether or not they believe the results to be reliable or trustworthy.

>>>>

# Extended Abstract

In data science, an analyst has created *reproducible* work if their code and data produces identical results when given to another analyst.

This idea of reproducibility is becoming increasingly important to the advancement of our scientific knowledge. One reason for this is that reproducible research encourages transparency. Providing one's audience with well organized, reproducible code and results allows readers to understand the steps taken to generate findings from raw data and determine for themselves whether or not they believe the results to be reliable or trustworthy.

Additionally, reproducibility encourages collaboration and extended research, allowing others to easily apply the methods used in one project to their own work with minimal effort.

However, while the necessity of reproducibility is clear, there are significant behavioral and technical challenges that impede its widespread implementation, and no clear consensus on standards of what constitutes reproducibility in published research.

• Note: What are these behavioral and technical challenges, and which ones should we mention?

In defining the steps toward creating reproducible analyses, researchers place their emphasis on different areas, including file structure, documentation, the use of file paths in code, and accessibility. Within these categories, the recommendations for what steps to take in ensuring reproducibility also vary.

• Why focus on R?

@@

One reason for this is that reproducible research encourages transparency. Providing one's audience with well organized, reproducible code and results allows readers to understand the steps taken to generate findings from raw data and determine for themselves whether or not they believe the results to be reliable or trustworthy.

### 2019-11-24 fertile.tex
start of extended abstract draft

\if0\blind
{
  \title{\bf Title here}

  \author{
        Author 1 \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of YYY, University of XXX\\
     and \\     Author 2 \\
    Department of ZZZ, University of WWW\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title here}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The text of your abstract. 200 or fewer words.
\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

@@

>>>>

\if0\blind
{
  \title{\bf Creating optimal conditions for reproducible data analysis in R with
\texttt{fertile}}

  \author{
        Audrey Bertin \thanks{The authors gratefully acknowledge contributions from Hadley Wickham,
Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the
\texttt{tidyverse} team.} \\
    Program in Statistical and Data Sciences, Smith College\\
     and \\     Benjamin S. Baumer \\
    Program in Statistical and Data Sciences, Smith College\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Creating optimal conditions for reproducible data analysis in R with
\texttt{fertile}}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
The advancement of scientific knowledge increasingly depends on ensuring
that data-driven research is reproducible: that two people with the same
data obtain the same results. However, while the necessity of
reproducibility is clear, there are significant behavioral and technical
challenges that impede its widespread implementation, and no clear
consensus on standards of what constitutes reproducibility in published
research. We focus on a series of common mistakes programmers make while
conducting data science projects in R, primarily through the RStudio
integrated development environment. \texttt{fertile} is an R package
that operates in two modes: proactively (to prevent reproducibility
mistakes from happening in the first place), and retroactively
(analyzing code that is already written for potential problems).
Furthermore, \texttt{fertile} is designed to educate the user about why
the mistakes are problematic, and how to fix them. We discuss
experimental results from testing \texttt{fertile} in an introductory
data science course.
\end{abstract}

\noindent%
{\it Keywords:} reproducibility, statistical software, workflow, collaboration
\vfill

@@

\section{Extended Abstract}\label{extended-abstract}

In data science, an analyst has created \emph{reproducible} work if
their code and data produces identical results when given to another
analyst.

This idea of reproducibility is becoming increasingly important to the
advancement of our scientific knowledge. One reason for this is that
reproducible research encourages transparency. Providing one's audience
with well organized, reproducible code and results allows readers to
understand the steps taken to generate findings from raw data and
determine for themselves whether or not they believe the results to be
reliable or trustworthy.

Additionally, reproducibility encourages collaboration and extended
research, allowing others to easily apply the methods used in one
project to their own work with minimal effort.

However, while the necessity of reproducibility is clear, there are
significant behavioral and technical challenges that impede its
widespread implementation, and no clear consensus on standards of what
constitutes reproducibility in published research.

• Note: What are these behavioral and technical challenges, and which
ones should we mention?

In defining the steps toward creating reproducible analyses, researchers
place their emphasis on different areas, including file structure,
documentation, the use of file paths in code, and accessibility. Within
these categories, the recommendations for what steps to take in ensuring
reproducibility also vary.

• Why focus on R?
